---
title: "Data Generation and Recommendation Engine Using Association Rules (Market Basket) in R"
date: '`r strftime(Sys.time(), "%Y-%m-%dT%H:%M:%S+03:00")`'
draft: false
description: ""
tags:
categories: recommendation, data
type: post
url:
author: "Mert Nuhoglu"
output: html_document
blog: datascience.mertnuhoglu.com
resource_files:
- version.csv
- schema.sql
- product_category.csv
- init.sql
- data.sql
- brand.csv
path: ~/projects/study/ds/recommendation_engine_using_association_rules_in_r/data_generation_for_recommendation_engine_using_association_rules_in_r.Rmd
---

<style>
  .main-container {
    max-width: 1600px !important;
  }
</style>

```{r}
library(dplyr, warn.conflicts = F)
library(dbplyr, warn.conflicts = F)
``` 

```{r}
df1 = readr::read_csv("brand.csv")
df2 = readr::read_csv("product_category.csv")
df3 = readr::read_csv("version.csv")
``` 

```{r}
df = merge(df1, df2, all = T) %>%
  merge(df3, all = T) %>%
  mutate( product_title = sprintf("%s %s %s", brand, product_category, version) %>% tolower ) 
writeLines(df$product_title, "product_title.txt")
``` 

``` {bash}
cat schema.sql
``` 

``` {bash}
datafiller --seed 1 --size=300 schema.sql > data.sql
``` 

``` {bash}
cat data.sql | sed -n "/COPY/,+2 p"
``` 

Now, we are going to start postgresql database server and import the generated data.

``` 
docker start postgreststarterkit_db_1
``` 

``` {bash}
psql -d app -h localhost -p 5432 -U superuser -f init.sql
``` 

Read data from PostgreSQL into R dataframe:

``` {r}
con2 <- DBI::dbConnect(RPostgreSQL::PostgreSQL()
  , user = Sys.getenv("SUPER_USER")
  , password = Sys.getenv("SUPER_USER_PASSWORD")
  , dbname = "app"
  , host = "localhost"
  , port = "5432"
  , options = "-c search_path=recommendation_engine"
)
prd = dplyr::tbl(con2, "product")
por = dplyr::tbl(con2, "purchase_order")
orl = dplyr::tbl(con2, "order_line")
```

Join the tables:

``` {r}
orl2 = orl %>%
  dplyr::inner_join(prd, by = "product_id") %>%
  dplyr::select(purchase_order_id, title, amount) %>%
  dplyr::arrange(purchase_order_id, title)
orl2
```

Target shape requires to concatenate all products in one purchase order transaction together:

``` {r}
orl3 = orl2 %>%
  dplyr::collect() %>%
  dplyr::group_by(purchase_order_id) %>%
  dplyr::mutate(key = sprintf("x%s", row_number())) %>%
  dplyr::ungroup() %>%
  dplyr::mutate( row = row_number() ) 
orl3

orl4 = orl3 %>%
  dplyr::select(purchase_order_id, key, title) %>%
  tidyr::spread(key = key, value = title) 
orl4 
```

```{r}
.libPaths("/Users/mertnuhoglu/.exploratory/R/3.4")
library(exploratory)

# Data Analysis Steps
res = orl4 %>%
  dplyr::select(-purchase_order_id) %>%
  dplyr::mutate(transaction_id = row_number()) %>%
  tidyr::gather(key, product, -transaction_id, na.rm = TRUE, convert = TRUE) %>%
  dplyr::arrange(transaction_id) %>%
  exploratory::do_apriori(product, transaction_id, min_support = 0.0001) %>%
  dplyr::filter(support > 0.0004) %>%
  dplyr::group_by(rhs) %>%
  dplyr::top_n(3, lift)
res
```

```{r}
res = orl4 %>%
  dplyr::select(-purchase_order_id) %>%
  dplyr::mutate(transaction_id = row_number()) %>%
  tidyr::gather(key, product, -transaction_id, na.rm = TRUE, convert = TRUE) %>%
  dplyr::arrange(transaction_id) %>%
  exploratory::do_apriori(product, transaction_id, min_support = 0.0001) %>%
  dplyr::filter(support > 0.0004) %>%
  dplyr::group_by(rhs) %>%
  dplyr::top_n(3, lift)
res
```

orl4 %>%
  dplyr::select(-purchase_order_id) %>%
  readr::write_csv("transactions_demo_data.csv")
res = orl4 %>%
  dplyr::select(-purchase_order_id) %>%
  dplyr::mutate(transaction_id = row_number()) %>%
  tidyr::gather(key, product, -transaction_id, na.rm = TRUE, convert = TRUE) %>%
  dplyr::arrange(transaction_id) 
