---
title: "R Date Time: Lubridate Examples"
output: html_document

---

# Article: 16 Dates and times | R for Data Science id=g12967

[16 Dates and times | R for Data Science](https://r4ds.had.co.nz/dates-and-times.html)

```{r}
library(lubridate)
library(tidyverse)
library(nycflights13)
```
## Creating date/times

3 types:

- date: printed as <date>
- time: printed as <time>
- date-time: printed as <dttm>. POSIXct ile denk.

Get current date or date-time: `today()` or `now()`:

```{r}
today()
  #> [1] "2022-06-23"
now()
  #> [1] "2022-06-23 22:37:37 +03"
```

3 ways to create a date/time:

- from a string
- from date-time components
- from an existing date/time object

### From string

a01: From quoted strings:
```{r}
ymd("2017-01-31")
  #> [1] "2017-01-31"
mdy("January 31st, 2017")
  #> [1] "2017-01-31"
dmy("31-Jan-2017")
  #> [1] "2017-01-31"
```

From unquoted numbers:

```{r}
ymd(20170131)
  #> [1] "2017-01-31"
```


`ymd()` creates a date.

`ymd_hms()` creates a date-time:

```{r}
ymd_hms("2017-01-31 20:11:59")
  #> [1] "2017-01-31 20:11:59 UTC"
mdy_hm("01/31/2017 08:01")
  #> [1] "2017-01-31 08:01:00 UTC"
ymd_hms("2017-01-31 20:11:59 +03")
  #> [1] "2017-01-31 17:11:59 UTC"
ymd_hms("2017-01-31 20:11:59 +00")
  #> [1] "2017-01-31 20:11:59 UTC"
ymd_hms("2017-01-31 20:11:59 +03", tz = "Turkey")
  #> [1] "2017-01-31 20:11:59 +03"

```

Timezone argument:

```{r}
ymd(20170131, tz = "UTC")
  #> [1] "2017-01-31 UTC"
```

### From individual components

```{r}
flights %>%
        select(year, month, day, hour, minute)
  #> # A tibble: 336,776 x 5
  #>    year month   day  hour minute
  #>   <int> <int> <int> <dbl>  <dbl>
  #> 1  2013     1     1     5     15
  #> 2  2013     1     1     5     29
```

Use `make_date()` for date, or `make_datetime()` for date-time:

```{r}
flights %>%
        select(year, month, day, hour, minute) %>%
        mutate(departure = make_datetime(year, month, day, hour, minute))
  #> # A tibble: 336,776 x 6
  #>    year month   day  hour minute departure
  #>   <int> <int> <int> <dbl>  <dbl> <dttm>
  #> 1  2013     1     1     5     15 2013-01-01 05:15:00
  #> 2  2013     1     1     5     29 2013-01-01 05:29:00
```

### From other types

Switch between date-time and date:
```{r}
as_datetime(today())
  #> [1] "2020-10-09 UTC"
as_date(now())
  #> [1] "2020-10-09"
```

From unix epoch:

```{r}
as_datetime(60 * 60 * 10)
  #> [1] "1970-01-01 10:00:00 UTC"
as_date(365 * 10 + 2)
  #> [1] "1980-01-01"
```

### Date-time components

Getting components: `year()`, `month()` ...

```{r}
datetime <- ymd_hms("2016-07-08 12:34:56")

year(datetime)
  #> [1] 2016
month(datetime)
  #> [1] 7
mday(datetime)
  #> [1] 8

yday(datetime)
  #> [1] 190
wday(datetime)
  #> [1] 6
```

`label = T` ile ayın ve günün ismini alırsın:

```{r}
month(datetime, label = TRUE)
  #> [1] Jul
wday(datetime, label = TRUE, abbr = FALSE)
  #> [1] Friday
```

Setting components in-place:

```{r}
(datetime <- ymd_hms("2016-07-08 12:34:56"))
  #> [1] "2016-07-08 12:34:56 UTC"

year(datetime) <- 2020
datetime
  #> [1] "2020-07-08 12:34:56 UTC"
month(datetime) <- 01
datetime
  #> [1] "2020-01-08 12:34:56 UTC"
hour(datetime) <- hour(datetime) + 1
datetime
  #> [1] "2020-01-08 13:34:56 UTC"
```

Setting components returning new objects: `update()`

```{r}
update(datetime, year = 2020, month = 2, mday = 2, hour = 2)
  #> [1] "2020-02-02 02:34:56 UTC"
```

## time spans

3 classes to represent time spans:

- duration: number of seconds
- period: human units like weeks and months
- intervals: starting and ending point

### Durations

Difftime object:
```{r}
  # How old is Hadley?
h_age <- today() - ymd(19791014)
h_age
  #> Time difference of 14971 days
```
Duation: always uses seconds

```{r}
as.duration(h_age)
  #> [1] "1293494400s (~40.99 years)"
```
Duration constructors:

```{r}
dseconds(15)
  #> [1] "15s"
dminutes(10)
  #> [1] "600s (~10 minutes)"
dhours(c(12, 24))
  #> [1] "43200s (~12 hours)" "86400s (~1 days)"
ddays(0:5)
  #> [1] "0s"                "86400s (~1 days)"  "172800s (~2 days)"
  #> [4] "259200s (~3 days)" "345600s (~4 days)" "432000s (~5 days)"
dweeks(3)
  #> [1] "1814400s (~3 weeks)"
dyears(1)
  #> [1] "31557600s (~1 years)"
```
Add and multiply durations:

```{r}
2 * dyears(1)
  #> [1] "63115200s (~2 years)"
dyears(1) + dweeks(12) + dhours(15)
  #> [1] "38869200s (~1.23 years)"
```
Add and subtract durations from dates or date-times:

```{r}
tomorrow <- today() + ddays(1)
last_year <- today() - dyears(1)
```

DST veya benzer özel durumlarda beklenmedik sonuçlar çıkabilir:

```{r}
one_pm <- ymd_hms("2016-03-12 13:00:00", tz = "America/New_York")

one_pm
  #> [1] "2016-03-12 13:00:00 EST"
one_pm + ddays(1)
  #> [1] "2016-03-13 14:00:00 EDT"

```
### Periods

İnsanın sezgisel zaman ölçü birimlerini temsil eder.

```{r}
one_pm
  #> [1] "2016-03-12 13:00:00 EST"
one_pm + days(1)
  #> [1] "2016-03-13 13:00:00 EDT"
```

Period constructors:

```{r}
seconds(15)
  #> [1] "15S"
minutes(10)
  #> [1] "10M 0S"
hours(c(12, 24))
  #> [1] "12H 0M 0S" "24H 0M 0S"
days(7)
  #> [1] "7d 0H 0M 0S"
months(1:6)
  #> [1] "1m 0d 0H 0M 0S" "2m 0d 0H 0M 0S" "3m 0d 0H 0M 0S" "4m 0d 0H 0M 0S"
  #> [5] "5m 0d 0H 0M 0S" "6m 0d 0H 0M 0S"
weeks(3)
  #> [1] "21d 0H 0M 0S"
years(1)
  #> [1] "1y 0m 0d 0H 0M 0S"
```

- Add and multiply periods
- Add them to dates

### Intervals

```{r}
years(1) / days(1)
  #> [1] 365.25
```

```{r}
next_year <- today() + years(1)
(today() %--% next_year) / ddays(1)
  #> [1] 365
```

```{r}
(today() %--% next_year) %/% days(1)
  #> [1] 365
```













