---
title: "Dplyr metaprogramming"
date: 2019-02-16T13:07:58+03:00 
draft: true
description: ""
tags:
categories: r dplyr metaprogramming
type: post
url:
author: "Mert Nuhoglu"
output: rmarkdown::html_document
blog: mertnuhoglu.com
resource_files:
-
---

<style>
  .main-container {
    max-width: 1600px !important;
  }
</style>

``` {r set-options}
options(width = 150)
options(max.print = 30)
``` 

## Article01: Programming with dplyr - vignette dplyr

Following https://cran.r-project.org/web/packages/dplyr/vignettes/programming.html

### Warm up

opt01: no parameter:

``` {r}
greet <- function(name) {
  "How do you do, name?"
}
greet("Hadley")
  ##> #> [1] "How do you do, name?"
``` 

opt02: parameterization with paste

``` {r}
greet <- function(name) {
  paste0("How do you do, ", name, "?")
}
greet("Hadley")
  ##> #> [1] "How do you do, Hadley?"
``` 

opt03: glue: unquote components

``` {r}
greet <- function(name) {
  glue::glue("How do you do, {name}?")
}
greet("Hadley")
  ##> #> How do you do, Hadley?
``` 

### Programming recipes

#### Different data sets

``` {r}
df <- tibble(
  g1 = c(1, 1, 2, 2, 2),
  g2 = c(1, 2, 1, 2, 1),
  a = sample(5),
  b = sample(5)
)
``` 

``` {r}
df %>%
  group_by(g1) %>%
  summarise(a = mean(a))
  ##> #> # A tibble: 2 x 2
  ##> #>      g1     a
  ##> #>   <dbl> <dbl>
  ##> #> 1     1  2.5 
  ##> #> 2     2  3.33
``` 

``` {r}
df %>%
  group_by(g2) %>%
  summarise(a = mean(a))
  ##> #> # A tibble: 2 x 2
  ##> #>      g2     a
  ##> #>   <dbl> <dbl>
  ##> #> 1     1   2  
  ##> #> 2     2   4.5
``` 

##### Goal: parameterize group_by arguments:

We have such a function:

``` r
my_summarise <- function(df, group_var) {
  df %>%
    group_by(group_var) %>%
    summarise(a = mean(a))
}
``` 

Call it like:

``` r
  ##> my_summarise(df, g1)
``` 

opt01: pass variable symbol

``` {r}

  ##> my_summarise(df, g1)
  ##> #> Error: Column `group_var` is unknown
``` 

opt02: pass string

``` {r}
  ##> my_summarise(df, "g2")
  ##> #> Error: Column `group_var` is unknown
``` 

opt03: quote variable

`quo`

``` {r}
quo(g1)
  ##> #> <quosure>
  ##> #> expr: ^g1
  ##> #> env:  global
quo(a + b + c)
  ##> #> <quosure>
  ##> #> expr: ^a + b + c
  ##> #> env:  global
quo("a")
  ##> #> <quosure>
  ##> #> expr: ^"a"
  ##> #> env:  empty
``` 

``` {r}
  ##> my_summarise(df, quo(g1))
  ##> #> Error: Column `group_var` is unknown
``` 

opt04: quote variable + tell group_by not to unquote it

``` {r}
my_summarise <- function(df, group_var) {
  df %>%
    group_by(!! group_var) %>%
    summarise(a = mean(a))
}

my_summarise(df, quo(g1))
  ##> #> # A tibble: 2 x 2
  ##> #>      g1     a
  ##> #>   <dbl> <dbl>
  ##> #> 1     1  2.5 
  ##> #> 2     2  3.33
``` 

But this still gives error:

``` {r}
  ##> my_summarise(df, quo(g1))
``` 

opt05: quote inside function

``` {r}
my_summarise <- function(df, group_var) {
  quo_group_var <- quo(group_var)
  print(quo_group_var)

  df %>%
    group_by(!! quo_group_var) %>%
    summarise(a = mean(a))
}

  ##> my_summarise(df, g1)
  ##> #> <quosure>
  ##> #> expr: ^group_var
  ##> #> env:  0x7ffe16901ea8
  ##> #> Error: Column `group_var` is unknown
``` 

`quo(group_var)` always returns `~group_var`. It is being too literal! We want it to substitute the value that the user supplied, i.e. to return `~g1`

opt06: evaluate lazily (promise)

By analogy to strings, we don’t want "", instead we want some function that turns an argument into a string. That’s the job of `enquo()`

``` {r}
my_summarise <- function(df, group_var) {
  group_var <- enquo(group_var)
  print(group_var)

  df %>%
    group_by(!! group_var) %>%
    summarise(a = mean(a))
}

my_summarise(df, g1)
  ##> #> <quosure>
  ##> #> expr: ^g1
  ##> #> env:  global
  ##> #> # A tibble: 2 x 2
  ##> #>      g1     a
  ##> #>   <dbl> <dbl>
  ##> #> 1     1  2.5 
  ##> #> 2     2  3.33
``` 

(If you’re familiar with `quote()` and `substitute()` in base R, `quo()` is equivalent to `quote()` and `enquo()` is equivalent to `substitute()`.)

#### Different input variable

Goal: make this a function:

``` {r}
summarise(df, mean = mean(a), sum = sum(a), n = n())
	#> # A tibble: 1 x 3
	#>    mean   sum     n
	#>   <dbl> <int> <int>
	#> 1     3    15     5
summarise(df, mean = mean(a * b), sum = sum(a * b), n = n())
	#> # A tibble: 1 x 3
	#>    mean   sum     n
	#>   <dbl> <int> <int>
	#> 1   9.6    48     5
``` 

opt01: `quo` the variable and unquote in dplyr with `!!`

``` {r}
my_var <- quo(a)
summarise(df, mean = mean(!! my_var), sum = sum(!! my_var), n = n())
	#> # A tibble: 1 x 3
	#>    mean   sum     n
	#>   <dbl> <int> <int>
	#> 1     3    15     5
``` 

Debugging dplyr: wrap dplyr call with `quo()`

``` {r}
quo(summarise(df,
  mean = mean(!! my_var),
  sum = sum(!! my_var),
  n = n()
))
	#> <quosure>
	#> expr: ^summarise(df, mean = mean(^a), sum = sum(^a), n = n())
	#> env:  global
``` 

opt02: `enquo` and `!!`

``` {r}
my_summarise2 <- function(df, expr) {
  expr <- enquo(expr)

  summarise(df,
    mean = mean(!! expr),
    sum = sum(!! expr),
    n = n()
  )
}
my_summarise2(df, a)
	#> # A tibble: 1 x 3
	#>    mean   sum     n
	#>   <dbl> <int> <int>
	#> 1     3    15     5
my_summarise2(df, a * b)
	#> # A tibble: 1 x 3
	#>    mean   sum     n
	#>   <dbl> <int> <int>
	#> 1   9.6    48     5
``` 

#### Different input and output variable

Now, the goal is to vary the name of the output variables:

``` r
mutate(df, mean_a = mean(a), sum_a = sum(a))
mutate(df, mean_b = mean(b), sum_b = sum(b))
``` 

``` r
my_mutate <- function(df, expr) {
  expr <- enquo(expr)
  mean_name <- paste0("mean_", quo_name(expr))
  sum_name <- paste0("sum_", quo_name(expr))

  mutate(df,
    !! mean_name := mean(!! expr),
    !! sum_name := sum(!! expr)
  )
}

my_mutate(df, a)
	#> # A tibble: 5 x 6
	#>      g1    g2     a     b mean_a sum_a
	#>   <dbl> <dbl> <int> <int>  <dbl> <int>
``` 

