---
title: "Build Data SQL: Insert and Update Statements"
date: '`r strftime(Sys.time(), "%Y-%m-%dT%H:%M:%S+03:00")`'
draft: false
description: ""
tags:
categories: r, sql
type: post
url:
author: "Mert Nuhoglu"
output: html_document
blog: mertnuhoglu.com
resource_files:
-
---

# Hello

<style>
  .main-container {
    max-width: 1600px !important;
  }
</style>

``` {r set-options}
options(width = 250)
options(max.print = 30)
``` 

```{r}
library(dplyr, warn.conflicts = F)
```

```{r}
evr = tibble::tribble(
  ~enum_var_id,     ~enum_var_name,
            1L,  "route_stop_enum",
            2L,      "action_enum",
            3L,       "place_enum"
  )
evl = tibble::tribble(
  ~enum_value_id,          ~enum_value_name,      ~enum_var_id,
           1000L,         "departure depot",                1L,
           1001L,            "return depot",                1L,
           1002L,               "pick stop",                1L,
           1003L,               "drop stop",                1L,
           1004L,          "pick/drop stop",                1L,
           1005L,  "departure/return depot",                1L,
           2000L,               "pick/load",                2L,
           2001L,             "drop/unload",                2L,
           3000L,                   "depot",                3L,
           3001L,                 "factory",                3L,
           3002L,                "customer",                3L
  )
```

Here are the test data:

``` {r}
den = tibble::tribble(
  ~data_entity_id,               ~entity_name,
               7L,               "enum_value",
               8L,                 "enum_var"
  )
dfl = tibble::tribble(
  ~data_field_id,         ~data_field_name,          ~type,     ~pk_fk, ~not_null, ~data_entity_id, ~fk_data_entity_id,     ~enum_var_name,
             26L,          "enum_value_id",          "INT",       "PK",     FALSE,              7L,                 NA,                 NA,
             27L,            "enum_var_id",          "INT",       "FK",      TRUE,              7L,                 8L,                 NA,
             28L,        "enum_value_name",         "TEXT",  "NON_KEY",     FALSE,              7L,                 NA,                 NA,
             29L,            "enum_var_id",          "INT",       "PK",     FALSE,              8L,                 NA,                 NA,
             30L,          "enum_var_name",         "TEXT",  "NON_KEY",     FALSE,              8L,                 NA,                 NA
  )
```

Of course, I didn't type this data manually. I reproduced this data using excellent life-saver library `datapasta` by running the following code on terminal once:

``` r
yumltordbschema::read_data_entity() %>% dplyr::as_data_frame() %>% datapasta::dpasta()
yumltordbschema::read_data_field() %>% dplyr::as_data_frame() %>% datapasta::dpasta()
```

`read_data_entity` and `read_data_field` functions import data from the file system. In order to make this `Rmd` file completely reproducible, I don't import data from the file system here.

The main entry function for sql generation is `rdb_to_data`:

``` r
rdb_to_data = function(entity, data_entities, dfl, den) {
  dflf = dfl %>%
    dplyr::left_join(den, by = "data_entity_id") %>%
    dplyr::filter(entity_name == entity) 
  columns = dflf$data_field_name
  id_column = dflf %>%
    dplyr::filter(pk_fk == "PK") %>%
    magrittr::extract2("data_field_name")
	df = data_entities[[entity]] %>%
    dplyr::select_(.dots = columns)
	build_sql(df, entity, id_column, den)
}
```

This will be called by the following lines in actual program:

``` r
data_model_dir = "/Users/mertnuhoglu/projects/itr/itr_documentation/data_model/"
r_entity = function(entity, data_model_dir = env_data_model_dir()) {
  rio::import(sprintf("%s/../rdb/%s.xlsx", data_model_dir, entity))
}
entities = c("enum_var", "enum_value")
data_entities = lapply(entities, r_entity, data_model_dir) %>%
	setNames(entities)
data_entities
lapply( entities, rdb_to_data, data_entities, dfl, den)
```

Note that, these statements access file system. In this Rmd file, I don't want to import data from file system. Instead I will initialize `data_entities` from sample dataframes:

``` {r}
entities = c("enum_var", "enum_value")
data_entities = list( enum_var = evr, enum_value = evl )
data_entities
``` 

Step by step, data flow inside `rdb_to_data()` is as follows:

``` {r}
entity = "enum_var"
dflf = dfl %>%
  dplyr::left_join(den, by = "data_entity_id") %>%
  dplyr::filter(entity_name == entity) 
columns = dflf$data_field_name
df = data_entities[[entity]] %>%
	dplyr::select_(.dots = columns)
df
```

So, we get meta data from `dfl` which stores all data fields for each entity (table). From that, we obtain the names of the columns (fields) which we will use while generating SQL commands.

Now, let's implement `build_sql()`

``` {r}
insert_sql_template = function( entity, columns ) {
	template = "INSERT INTO %s (%s) VALUES (%s);"
	column_names = columns %>% paste(collapse=",")
	data_placeholders = rep( "'%s'", length(columns ) ) %>% paste(collapse=",") 
	result = sprintf( template, entity, column_names, data_placeholders )
	return(result)
}

insert_sql = function( data, template ) {
	arg = c( list(template), as.list(data) )
	do.call( 'sprintf', arg ) %>%
		stringr::str_replace_all( "'NA'", "null" )
}

sql_insert = function(df, entity) {
	template_df = dplyr::data_frame( entity = entity, insert_template = insert_sql_template(entity, names(df) ))
	insert_sql( df, template_df$insert_template )
}

sql = sql_insert( df, entity )
sql
```

`df` contains the data for which we generate `SQL INSERT` commands.

`sql_insert()` function converts data contained in a `data_frame` into `SQL INSERT` statements.

``` {r}
build_sql = function(df, entity) {
	sql = sql_insert( df, entity )
  return(list(
              sql_insert = sql
              ))
}
``` 

I put `sql_insert` inside `build_sql` function because later I will also create `UPDATE` and other commands in addition to `INSERT` inside `build_sql()`.

So, let's update definition of `rdb_to_data()` with this `build_sql()` and run the program again:

``` {r}
rdb_to_data = function(entity, data_entities, dfl, den) {
  dflf = dfl %>%
    dplyr::left_join(den, by = "data_entity_id") %>%
    dplyr::filter(entity_name == entity) 
  columns = dflf$data_field_name
	df = data_entities[[entity]] %>%
    dplyr::select_(.dots = columns)
	build_sql(df, entity)
}
lapply( entities, rdb_to_data, data_entities, dfl, den)
```

Now, we can define functions that will generate `SQL UPDATE` and `SQL DELETE` commands. The functions are very similar to `SQL INSERT` generating functions.

Let's add `SQL UPDATE` and `SQL DELETE` statements into `build_sql()` function:

``` {r}
delete_sql = function( df, template ) {
	arg = c( list(template), as.list(df) )
	do.call( 'sprintf', arg ) 
}

delete_sql_template = function( entity, id_column) {
	template = "DELETE FROM %s WHERE %s = %%s;"
	sprintf( template, entity, id_column)
}

sql_delete = function(df, entity) {
	template = dplyr::data_frame( entity= entity, delete_template = delete_sql_template(entity, names(df) ))
	delete_sql( df, template$delete_template )
}

update_sql = function( data, template ) {
	arg = c( list(template), as.list(data) )
	do.call( 'sprintf', arg ) %>%
		stringr::str_replace_all( "'\\bNA\\b'", "null" ) 
}

update_sql_template = function( entity, columns ) {
	template = "    UPDATE %s SET %s WHERE %s = %%s;"
	columns_to_set = head(columns, -1)
	set_column_clause = sprintf("%s = '%%s'", columns_to_set) %>% 
		paste(collapse=", ")
  id_column = tail(columns, 1)
	sprintf( template, entity, set_column_clause, id_column )
}

sql_update = function(df, entity) {
	template = dplyr::data_frame( entity= entity, update_template = update_sql_template(entity, names(df) ))
	update_sql( df, template$update_template )
}
```

Now, let's update `build_sql` with `sql_update`, `sql_delete`:

``` {r}
build_sql = function(df, entity, id_column, den) {
  df_id_at_end = df %>%
		dplyr::select( -dplyr::one_of(id_column), dplyr::everything(), id_column )
  df_no_fk = dplyr::select( df, -dplyr::ends_with("_id"), id_column) %>%
    dplyr::select( id_column, dplyr::everything() )
  if ( any("invalid" %in% names(df)) ) {
    df_invalid = df %>%
      dplyr::filter( invalid == 1 ) %>%
      dplyr::select(id_column)
  } else {
    df_invalid = df %>%
      dplyr::select(id_column)
  }
  list(
       sql_insert = sql_insert( df, entity )
       , sql_update = sql_update( df_id_at_end, entity )
       , sql_insert_no_fk = sql_insert( df_no_fk, entity )
       , sql_delete = sql_delete( df_invalid, entity )
       )
} 

rdb_to_data = function(entity, data_entities, dfl, den) {
  dflf = dfl %>%
    dplyr::left_join(den, by = "data_entity_id") %>%
    dplyr::filter(entity_name == entity) 
  columns = dflf$data_field_name
  id_column = dflf %>%
    dplyr::filter(pk_fk == "PK") %>%
    magrittr::extract2("data_field_name")
	df = data_entities[[entity]] %>%
    dplyr::select_(.dots = columns)
	build_sql(df, entity, id_column, den)
}
ent_m_cmdtype_m_cmds = lapply( entities, rdb_to_data, data_entities, dfl, den) %>%
  setNames(entities)
str(ent_m_cmdtype_m_cmds)
```

## The Essence of SQL Generation

The essence of SQL commands generation are done by the following functions:

``` r
insert_sql_template = function( entity, columns ) {
	template = "INSERT INTO %s (%s) VALUES (%s);"
	column_names = columns %>% paste(collapse=",")
	data_placeholders = rep( "'%s'", length(columns ) ) %>% paste(collapse=",") 
	result = sprintf( template, entity, column_names, data_placeholders )
	return(result)
}

insert_sql = function( data, template ) {
	arg = c( list(template), as.list(data) )
	do.call( 'sprintf', arg ) %>%
		stringr::str_replace_all( "'NA'", "null" )
}

sql_insert = function(df, entity) {
	template_df = dplyr::data_frame( entity = entity, insert_template = insert_sql_template(entity, names(df) ))
	insert_sql( df, template_df$insert_template )
}

sql = sql_insert( df, entity )
```

`insert_sql_template()` function returns a template string:

``` {r}
template_df = dplyr::data_frame( entity = entity, insert_template = insert_sql_template(entity, names(df) ))
template_df
```

This template string is filled with the data in `df`. 

``` {r}
df
```

