---
title: "Study testthat unit tests"
date: 2019-06-27T15:09:26+03:00 
draft: true
description: ""
tags:
categories: r statistics
type: post
url:
author: "Mert Nuhoglu"
output: rmarkdown::html_document
blog: mertnuhoglu.com
resource_files:
-
---

<style>
  .main-container {
    max-width: 1600px !important;
  }
</style>

``` {r set-options}
options(width = 150)
options(max.print = 30)
``` 

## Article01: 

https://kbroman.org/pkg_primer/pages/tests.html

## Article02: 

http://r-pkgs.had.co.nz/tests.html

First create a project

```{r}
library(usethis)

create_package("ex02")
``` 

```{r}
usethis::use_testthat()
usethis::use_test("test01")
  ##> ~/projects/study/r/ex/study_testthat_unit_tests/ex02/tests/testthat/test-test01.R
``` 

Run tests

```{r}
devtools::test()
``` 

## Article03: 

https://www.r-bloggers.com/unit-testing-shiny-apps-using-testthat/

``` bash
cd ~/projects/study/r/ex/study_testthat_unit_tests
R
``` 

```{r}
library(usethis)

create_package("art03")
usethis::use_testthat()
use_package("shiny")
``` 

Edit `~/projects/study/r/ex/study_testthat_unit_tests/art03/R/app.R`

```{r}
source("R/app.R")
app = shiny::shinyApp(ui, server)
runApp(app)
``` 

# Examples

## ex03: use package functions in test script

``` bash
cd ~/projects/study/r/ex/study_testthat_unit_tests
R
``` 

```{r}
library(usethis)

create_package("ex03")
``` 

```{r}
usethis::use_testthat()
usethis::use_test("test01")
  ##> ~/projects/study/r/ex/study_testthat_unit_tests/ex03/tests/testthat/test-test01.R
``` 

Edit `~/projects/study/r/ex/study_testthat_unit_tests/ex03/R/funs.R`

```{r}
ex03_fun = function() {
	print("hello")
}
``` 

Edit `~/projects/study/r/ex/study_testthat_unit_tests/ex03/tests/testthat/test-test01.R`

```{r}
test_that("can use library functions", {
	ex03_fun()
  expect_equal(1, 1)
})
``` 

Run tests

```{r}
devtools::test()
``` 

All tests pass.

## ex03: use package functions in test script



