---
title: "Study stringr
author: "Mert Nuhoglu"
output:
  html_document: default
    toc: true
    toc_float: true

---

```{r load_packages, include=FALSE}
    source("utils_study.R")
    ss()
```

## sed s: str_replace str_replace_all

```{r}
str_replace(string, pattern, replacement)
``` 

```{r}
library(stringr)
fruits <- c("one apple", "two pears", "three bananas")
str_replace(fruits, "[aeiou]", "-")
  ##> [1] "-ne apple"     "tw- pears"     "thr-e bananas"
str_replace_all(fruits, "[aeiou]", "-")
  ##> [1] "-n- -ppl-"     "tw- p--rs"     "thr-- b-n-n-s"
str_replace_all(fruits, "[aeiou]", toupper)
  ##> [1] "OnE ApplE"     "twO pEArs"     "thrEE bAnAnAs"
str_replace_all(fruits, "b", NA_character_)
  ##> [1] "one apple" "two pears" NA
```

### str_match

```{r}
str_match(string, pattern)
str_match_all(string, pattern)
``` 

```{r}
library(stringr)
strings = c(" 219 733 8965", "329-293-8753 ", "banana")
pattern <- "([2-9][0-9]{2})[- .]([0-9]{3})[- .]([0-9]{4})"
str_extract(strings, pattern)
  ##> [1] "219 733 8965" "329-293-8753" NA
m = str_match(strings, pattern)
m
  ##>      [,1]           [,2]  [,3]  [,4]  
  ##> [1,] "219 733 8965" "219" "733" "8965"
  ##> [2,] "329-293-8753" "329" "293" "8753"
  ##> [3,] NA             NA    NA    NA
m[1,1] # match 1 group 1
  ##> [1] "219 733 8965"
m[1,2] # match 1 group 2
  ##> [1] "219"
```

## grep grepl subset: str_subset str_which str_detect

https://stringr.tidyverse.org/reference/str_subset.html
 
`str_subset()` is a wrapper around `x[str_detect(x, pattern)]`, and is equivalent to `grep(pattern, x, value = TRUE)`. `str_which()` is a wrapper around `which(str_detect(x, pattern))`, and is equivalent to `grep(pattern, x)`. See `str_detect()` for an equivalent to `grepl(pattern, x)`.

```{r}
	str_subset(string, pattern, negate = FALSE)
	str_which(string, pattern, negate = FALSE)
``` 

### Examples

```{r}
fruit <- c("apple", "banana", "pear", "pinapple")
str_subset(fruit, "a")
  #> [1] "apple"    "banana"   "pear"     "pinapple"
str_which(fruit, "a")
  #> [1] 1 2 3 4
str_subset(fruit, "^a")
  #> [1] "apple"
str_subset(fruit, "a$")
  #> [1] "banana"
str_subset(fruit, "b")
  #> [1] "banana"
str_subset(fruit, "[aeiou]")
  #> [1] "apple"    "banana"   "pear"     "pinapple"
  # Returns elements that do NOT match
str_subset(fruit, "^p", negate = TRUE)
  #> [1] "apple"  "banana"
  # Missings never match
str_subset(c("a", NA, "b"), ".")
  #> [1] "a" "b"
str_which(c("a", NA, "b"), ".")
  #> [1] 1 3
``` 

## detect

https://stringr.tidyverse.org/reference/str_detect.html

Equivalent to `grepl(pattern, x)`. See `str_which()` for an equivalent to `grep(pattern, x)`.

### Examples

```{r}
fruit <- c("apple", "banana", "pear", "pinapple")

str_detect(fruit, "a")
  ##> [1] TRUE TRUE TRUE TRUE
str_detect(fruit, "^a")
  ##> [1]  TRUE FALSE FALSE FALSE
str_detect(fruit, "a$")
  ##> [1] FALSE  TRUE FALSE FALSE
str_detect(fruit, "b")
  ##> [1] FALSE  TRUE FALSE FALSE
str_detect(fruit, "[aeiou]")
  ##> [1] TRUE TRUE TRUE TRUE

# Also vectorised over pattern

str_detect("aecfg", letters)
  ##>  [1]  TRUE FALSE  TRUE FALSE  TRUE  TRUE  TRUE FALSE FALSE FALSE FALSE FALSE

  ##> [13] FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE

  ##> [25] FALSE FALSE

# Returns TRUE if the pattern do NOT match

str_detect(fruit, "^p", negate = TRUE)
  ##> [1]  TRUE  TRUE FALSE FALSE
``` 




