---
title: "Studying R"
date: 2018-04-14T16:06:01+03:00 
draft: false
description: ""
tags:
categories: r
type: post
url:
author: "Mert Nuhoglu"
output:
  html_document:
    css: styles.css
state: wip

---

```{r load_packages, include=FALSE}
    source("utils_study.R")
    ss()
```

## Giriş

- http://mertnuhoglu.github.io/data_analysis_presentations/r_verigazeteciligi/
- Örneklere dayalı dil öğrenme

```{r }
	 library(data.table)
	 flights = fread("flights.csv")
	 airports = fread("airports.csv")
	 planes = fread("planes.csv")
	 weather = fread("weather.csv")
	 str(weather)
```

## Example 01

```{r }
  study_remove_all_na_columns = function() {
    df = data.frame( a = c(1,2), b = NA, c = NA )
    # find all na columns
    # opt1: works
    apply(is.na(df), 2, all)
    # opt2: wrong
    lapply(is.na(df), all)
    # opt3: works
    lapply(df, function(col) all(is.na(col)) )
    # opt4: works
    lapply(df, . %>% is.na %>% all )
    # filtering columns
    # opt1: correct
    df[ c(T, F, F) ]
    # opt2: correct
    df[ lapply(df, . %>% is.na %>% all %>% not ) %>% unlist ]
    # opt3: correct
    df[ apply(!is.na(df), 2, all) ]
    # opt4: correct
    df[ applyc(!is.na(df), all) ] }
```

## Ex02: Matrix to dataframe


``` bash
m0 = matrix(c("a","b","c",1,2,3,4,5,6,7,8,9),ncol = 3)
titles = m0[1,]
d0 = dplyr::as_data_frame(m0)
setNames(d0, titles)
  ##> # A tibble: 4 x 3
  ##>   a     `2`   `6`
  ##>   <chr> <chr> <chr>
  ##> 1 a     2     6
  ##> 2 b     3     7
  ##> 3 c     4     8
  ##> 4 1     5     9
``` 

## Ex03:

``` bash
sevk_emri_data = jsonlite::fromJSON('{"sevk_emri_data": [["irsaliye_no", "irsaliye_sevk_tarihi", "cari_kod", "unvan", "teslIm_adres", "teslim_sehir", "teslim_sehir__1", "malzeme_kodu", "malzeme_adi", "irsaliye_KG", "tesis", "IL", "ILCE", "Yaş/Toz"], ["1043", "2017-01-02T00:00:00Z", "2I00048", "ISIKSAN YALITIM INS. TAAH. SAN. VE TIC. LTD. STI.", "FABRIKADAN TESLIM ALACAGIZ.", "TEKİRDAĞ", "TÜRKİYE", "1032248PX20", "PE 5088 RAL3004 Smooth", "20", "ÇERKEZKÖY", "TEKIRDAG", "CERKEZKOY", "Toz"], ["1044", "2017-01-02T00:00:00Z", "1P00009", "BSH EV ALETLERI SANAYI VE TICARET A.S.", "YILDIRIM BEYAZID MAH.HAYRI BATUR CAD. NO: 72 CERKEZKOY", "TEKİRDAĞ", "TÜRKİYE", "1021378EB500", "G Miles+ W87 Vzf 07020 Smooth", "1000", "ÇERKEZKÖY", "TEKIRDAG", "CERKEZKOY", "Toz"], ["1045", "2017-01-02T00:00:00Z", "1P00009", "BSH EV ALETLERI SANAYI VE TICARET A.S.", "YILDIRIM BEYAZID MAH.HAYRI BATUR CAD. NO: 72 CERKEZKOY", "TEKİRDAĞ", "TÜRKİYE", "1023813EB300", "G Miles+ W88 Vzf 07020 Smooth", "300", "ÇERKEZKÖY", "TEKIRDAG", "CERKEZKOY", "Toz"]]}')$sevk_emri_data
titles = sevk_emri_data[1,]
sem = dplyr::as_data_frame(sevk_emri_data[2:nrow(sevk_emri_data),]) %>%
	setNames(titles)
``` 

## Ex04: sprintf with multiple vectors

``` R
sprintf("%s %s", "a", "a")
  ##> [1] "a a"
sprintf("%s %s", "a", c("a", "b"))
  ##> [1] "a a" "a b"
``` 

I want to obtain:

		[1] "a a" "a b" 
		->
		[1] "a a b"

``` R
do.call( 'sprintf', list("a", "a", "b") ) 
  ##> [1] "a"
do.call( 'sprintf', as.list(c("a", "a", "b")) ) 
  ##> [1] "a"
list("a", "a", "b")
  ##> [[1]]
  ##> [1] "a"
  ##> 
  ##> [[2]]
  ##> [1] "a"
  ##> 
  ##> [[3]]
  ##> [1] "b"
``` 

``` R
sprintf("%s %s", c("a", "a", "b"))
  ##> Error in sprintf("%s %s", c("a", "a", "b")) : too few arguments
sprintf("min %s", c(1, 2, 3))
  ##> [1] "min 1" "min 2" "min 3"
``` 

Prepare the argument before sprintf:

``` R
sprintf("%s %s", "a", paste("a", "b"))
  ##> [1] "a a b"
``` 

## Ex05: sprintf with very long arg

``` R
Error in sprintf(sql04, "data_20180903", "data_20181202") :
  'fmt' length exceeds maximal format length 8192
``` 

Solution:

Use paste0

## Ex06: sprintf errors

``` sql
columns = c(
	"bq_company_legal_name",
	"bq_company_commonly_known_as_name",
	"company_ein",
	"bq_ticker",
	"bq_cusip",
	"bq_cik",
	"bq_website",
	"bq_company_contact_name",
	"bq_company_address1_line_1",
	"bq_company_address1_line_2",
	"bq_company_address1_city",
	"bq_company_address1_state",
	"bq_company_address1_zip5",
	"bq_comp_fgn_address1_cntry",
	"bq_company_phone",
	"bq_age_oldest_plan",
	"bq_business_code",
	"bq_industry_name",
	"bq_industry_sub_sector",
	"bq_industry_sector",
	"bq_naics_code",
	"bq_naics_description",
	"bq_sic_code",
	"bq_sic_description",
	"bq_gics_code",
	"bq_gics_description",
	"bq_year",
	"bq_comp_most_recent_filing_date",
	"bq_age_oldest_pens_plan_p",
	"bq_age_oldest_welf_plan_p",
	"bq_mths_since_latest_plan_file_p",
	"bq_age_oldest_pens_plan",
	"bq_age_oldest_welf_plan",
	"bq_mths_since_latest_plan_file",
	"bq_data_run_date",
	"bq_no_filings_prev_2_yrs_ind",
	"bq_closed_all_plans_ind",
	"bq_public_indicator",
	"bq_small_company_ind"
)
sprintf("distinct(cnt, company_ein, %s) %>% nrow()", columns)
do.call( sprintf, c( fmt = "distinct(cnt, company_ein, %s) %>% nrow()", as.list(columns)) ) 
``` 

``` R
lapply(columns, function(v) 
sprintf("distinct(cnt, company_ein, %s) %>% nrow()", v)
sprintf("%s %>%", v)
sprintf("%s) %>%", v)
sprintf("distinct(cnt, company_ein, %s", v)
sprintf("distinct(cnt, company_ein, %s)", v)
sprintf("distinct(cnt, company_ein, %s) %>%", v)
sprintf("distinct(cnt, company_ein, %s) %>% nrow()", v)
)
``` 

## Ex07: sample function

``` r
x = letters[1:5]
set.seed(1)
sample(x)
  #> [1] "b" "e" "d" "c" "a"
sample(x, replace = F)
  #> [1] "e" "d" "b" "c" "a"
sample(x, replace = T)
  #> [1] "b" "a" "d" "b" "d"
sample(x, size = 3, replace = T)
  #> [1] "c" "d" "e"
```

## Ex08: probability distributions

```{r}
b0 = rbeta(5, 6, 3)
b0
mean(b0)
```

## Ex09: Sample quantiles

```{r}
q0 = quantile(x = rnorm(100))
q0
str(q0)
```

```{r}
quantile(x = rnorm(10), probs = seq(0, 1, 0.1))
```

## Ex10: rep function

``` {r}
rep.int(1:3, 2)
  ##> [1] 1 2 3 1 2 3
rep(2, 3)
  ##> [1] 2 2 2
rep(1:3, 2)
  ##> [1] 1 2 3 1 2 3
rep(1:3, each = 2)
  ##> [1] 1 1 2 2 3 3
rep(1:3, c(2,2,2))
  ##> [1] 1 1 2 2 3 3
rep(1:3, c(2,1,1))
  ##> [1] 1 1 2 3
rep(1:3, each = 2, len = 4)
  ##> [1] 1 1 2 2
rep(1:3, each = 2, times = 3)
  ##>  [1] 1 1 2 2 3 3 1 1 2 2 3 3 1 1 2 2 3 3
``` 

``` {r}
xs = list(h = 1:3, name = "a")
xs
  ##> $h
  ##> [1] 1 2 3
  ##> 
  ##> $name
  ##> [1] "a"
rep(xs, 2)
  ##> $h
  ##> [1] 1 2 3
  ##> 
  ##> $name
  ##> [1] "a"
  ##> 
  ##> $h
  ##> [1] 1 2 3
  ##> 
  ##> $name
  ##> [1] "a"
``` 

## Ex11: seq function

``` {r}
seq(0, 1, length.out = 3)
  ##> [1] 0.0 0.5 1.0
seq(stats::runif(3))
  ##> [1] 1 2 3 
seq(stats::runif(5))
  ##> [1] 1 2 3 4 5
seq(stats::rnorm(5))
  ##> [1] 1 2 3 4 5
seq(1, 7, by = 2)
  ##> [1] 1 3 5 7
seq(1, 7, by = pi)
  ##> [1] 1.000000 4.141593
``` 

## Ex12: paste function to concat/join lines

https://stat.ethz.ch/R-manual/R-devel/RHOME/library/base/html/paste.html

``` r
paste0(1:12)
  ##>  [1] "1"  "2"  "3"  "4"  "5"  "6"  "7"  "8"  "9"  "10" "11" "12"
paste(1:12)        # same
  ##>  [1] "1"  "2"  "3"  "4"  "5"  "6"  "7"  "8"  "9"  "10" "11" "12"
as.character(1:12) # same
  ##>  [1] "1"  "2"  "3"  "4"  "5"  "6"  "7"  "8"  "9"  "10" "11" "12"

  ## If you pass several vectors to paste0, they are concatenated in a
  ## vectorized way.
(nth <- paste0(1:12, c("st", "nd", "rd", rep("th", 9))))
  ##>  [1] "1st"  "2nd"  "3rd"  "4th"  "5th"  "6th"  "7th"  "8th"  "9th"  "10th"
  ##> [11] "11th" "12th"

  ## paste works the same, but separates each input with a space.
  ## Notice that the recycling rules make every input as long as the longest input.
paste(month.abb, "is the", nth, "month of the year.")
  ##>  [1] "Jan is the 1st month of the year." 
  ##>  [2] "Feb is the 2nd month of the year." 
  ##>  [3] "Mar is the 3rd month of the year." 
  ##>  [4] "Apr is the 4th month of the year." 
  ##>  [5] "May is the 5th month of the year." 
  ##>  [6] "Jun is the 6th month of the year." 
  ##>  [7] "Jul is the 7th month of the year." 
  ##>  [8] "Aug is the 8th month of the year." 
  ##>  [9] "Sep is the 9th month of the year." 
  ##> [10] "Oct is the 10th month of the year."
  ##> [11] "Nov is the 11th month of the year."
  ##> [12] "Dec is the 12th month of the year."
paste(month.abb, letters)
  ##>  [1] "Jan a" "Feb b" "Mar c" "Apr d" "May e" "Jun f" "Jul g" "Aug h"
  ##>  [9] "Sep i" "Oct j" "Nov k" "Dec l" "Jan m" "Feb n" "Mar o" "Apr p"
  ##> [17] "May q" "Jun r" "Jul s" "Aug t" "Sep u" "Oct v" "Nov w" "Dec x"
  ##> [25] "Jan y" "Feb z"

  ## You can change the separator by passing a sep argument
  ## which can be multiple characters.
paste(month.abb, "is the", nth, "month of the year.", sep = "_*_")
  ##>  [1] "Jan_*_is the_*_1st_*_month of the year." 
  ##>  [2] "Feb_*_is the_*_2nd_*_month of the year." 
  ##>  [3] "Mar_*_is the_*_3rd_*_month of the year." 
  ##>  [4] "Apr_*_is the_*_4th_*_month of the year." 
  ##>  [5] "May_*_is the_*_5th_*_month of the year." 
  ##>  [6] "Jun_*_is the_*_6th_*_month of the year." 
  ##>  [7] "Jul_*_is the_*_7th_*_month of the year." 
  ##>  [8] "Aug_*_is the_*_8th_*_month of the year." 
  ##>  [9] "Sep_*_is the_*_9th_*_month of the year." 
  ##> [10] "Oct_*_is the_*_10th_*_month of the year."
  ##> [11] "Nov_*_is the_*_11th_*_month of the year."
  ##> [12] "Dec_*_is the_*_12th_*_month of the year."

  ## To collapse the output into a single string, pass a collapse argument.
paste0(nth, collapse = ", ")
  ##> [1] "1st, 2nd, 3rd, 4th, 5th, 6th, 7th, 8th, 9th, 10th, 11th, 12th"

  ## For inputs of length 1, use the sep argument rather than collapse
paste("1st", "2nd", "3rd", collapse = ", ") # probably not what you wanted
  ##> [1] "1st 2nd 3rd"
paste("1st", "2nd", "3rd", sep = ", ")
  ##> [1] "1st, 2nd, 3rd"

  ## You can combine the sep and collapse arguments together.
paste(month.abb, nth, sep = ": ", collapse = "; ")
  ##> [1] "Jan: 1st; Feb: 2nd; Mar: 3rd; Apr: 4th; May: 5th; Jun: 6th; Jul: 7th; Aug: 8th; Sep: 9th; Oct: 10th; Nov: 11th; Dec: 12th"
Created on 2019-07-28 by the reprex package (v0.2.1)
``` 


## Article01: Generating Data in R

https://labs.uswitch.com/generating-data-in-r/

```{r}
  #generate 5 samples 
runif(5, min = 0, max = 10) 
```

```{r}
  #We can extend the runif() function like this:
runif_vec_proto <- function(min, max) {return(runif(1, min = min, max = max))}  
runif_vec <- Vectorize(runif_vec_proto); runif_vec_proto <- NULL

  # uniform dists 0-5 3-6 3-6 7-10. take one sample from each
runif_vec(min = c(0, 3, 3, 7), max = c(5 , 6, 6, 10))  
```

### Ex01: Probability that B is better than A

```{r}
observations_in_A <- 5000; observations_in_B <- 5000;  
successes_in_A <- 510; successes_in_B <- 553

successes_in_A/observations_in_A;successes_in_B/observations_in_B;
```

```{r}
  #fails calculations for beta distribution
fails_in_A <- observations_in_A - successes_in_A  
fails_in_B <- observations_in_B - successes_in_B

  #Sample from the posterior distribution for each variant, assuming uninformative prior:
samples <- 1000000  
sampled_true_success_rates_A <- rbeta(samples, successes_in_A+1, fails_in_A+1)  
sampled_true_success_rates_B <- rbeta(samples, successes_in_B+1, fails_in_B+1)  
sampled_absolute_increase_in_BvsA <- sampled_true_success_rates_B-sampled_true_success_rates_A

  #Look at the quantiles, for interest
quantile(sampled_absolute_increase_in_BvsA, seq(0,1, 0.05))

  #Estimate the chance that the true success rate of B is greater than that of A
mean(sampled_absolute_increase_in_BvsA>0)  
```

### Ex03: Compare Logistic regression with Neural Network

```{r}
  #install.packages(c(‘dplyr’, ‘nnet’, ‘ggplot2’, 'reshape2'))
require(dplyr); require(nnet); require(ggplot2); require(reshape2)

  #generate independent variable samples
x <- runif(1000, 0, 1000)

  #generate success/fail observations, which has a probability which follows 
  #a logistic curve which is close to 0 around x = 0, 
  #close to 0.6 around x=1000, and close to 0.3 around x=500
d <- 0.6*plogis(x, location = 500, scale = 30)  
rbinom_vec_proto <- function(p, size = 1) {return(rbinom(1, size, p))}  
rbinom_vec <- Vectorize(rbinom_vec_proto); rbinom_vec_proto <- NULL  
y <- rbinom_vec(d)

  #Add data to data frame
data <- data.frame(x = x,  y = y)

  #Add the fitted values from standard logistic regression and from a single neuron neural network to the data frame
m <- glm(y~x, data, family = binomial())  
f <- m$fitted.values  
m2 <- nnet(y~x, data, size = 1, entropy = TRUE, abstol = 1.0e-7, reltol = 1.0e-10)  
f2 <- m2$fitted.values  
data  <- data.frame(x = x, y = y, p = d, logistic = f, nnet = f2)

  #Plot the original outcome probability, the binary outcome data, 
  #and the fitted values from our two models for each x:
data %>% melt('x') %>% ggplot(aes(x, value, color = variable)) + geom_point()  
```

### purrr::splice id=g_10796

Follows https://www.rdocumentation.org/packages/purrr/versions/0.2.5/topics/splice

``` r
> library(purrr)
Warning message:
package ‘purrr’ was built under R version 3.5.2
> inputs <- list(arg1 = "a", arg2 = "b")
>
> # splice() concatenates the elements of inputs with arg3
> splice(inputs, arg3 = c("c1", "c2")) %>% str()
List of 3
 $ arg1: chr "a"
 $ arg2: chr "b"
 $ arg3: chr [1:2] "c1" "c2"
> list(inputs, arg3 = c("c1", "c2")) %>% str()
List of 2
  ..$ arg1: chr "a"
  ..$ arg2: chr "b"
 $ arg3: chr [1:2] "c1" "c2"
> c(inputs, arg3 = c("c1", "c2")) %>% str()
List of 4
 $ arg1 : chr "a"
 $ arg2 : chr "b"
 $ arg31: chr "c1"
 $ arg32: chr "c2"
``` 

### Article: Contingency tables in R

Follows https://www.datacamp.com/community/tutorials/contingency-tables-r

#### How to make a table

``` r
library(MASS)
Cars93$Type
table(Cars93$Type)
  ##> Compact   Large Midsize   Small  Sporty     Van 
  ##>      16      11      22      21      14       9 
``` 

``` r
prop.table(table(Cars93$Type))
``` 

#### How to make a contingency table

Using two frequencies combined:

``` r
table(Cars93$Type)
  ##> Compact   Large Midsize   Small  Sporty     Van 
  ##>      16      11      22      21      14       9 
table(Cars93$Origin)
  ##>     USA non-USA
  ##>      48      45
table(Cars93$Type, Cars93$Origin)
  ##>           USA non-USA
  ##>   Compact   7       9
  ##>   Large    11       0
  ##>   Midsize  10      12
  ##>   Small     7      14
  ##>   Sporty    8       6
  ##>   Van       5       4
str(table(Cars93$Type, Cars93$Origin))
  ##>  'table' int [1:6, 1:2] 7 11 10 7 8 5 9 0 12 14 ...
``` 

#### Marginals from contingency table: row sums

``` r
t0 = table(Cars93$Type, Cars93$Origin)
rowSums(t0)
  ##> Compact   Large Midsize   Small  Sporty     Van
  ##>      16      11      22      21      14       9
colSums(t0)
``` 

#### Percents from contingency table

Percentages overall:

``` r
t1 = prop.table(table(Cars93$Type, Cars93$Origin))
t1
  ##>                  USA    non-USA
  ##>   Compact 0.07526882 0.09677419
  ##>   Large   0.11827957 0.00000000
  ##>   Midsize 0.10752688 0.12903226
  ##>   Small   0.07526882 0.15053763
  ##>   Sporty  0.08602151 0.06451613
  ##>   Van     0.05376344 0.04301075
``` 

Percentages row-wise:

``` r
t2 = prop.table(table(Cars93$Type, Cars93$Origin), margin=1)
t2
  ##>                 USA   non-USA
  ##>   Compact 0.4375000 0.5625000
  ##>   Large   1.0000000 0.0000000
  ##>   Midsize 0.4545455 0.5454545
  ##>   Small   0.3333333 0.6666667
  ##>   Sporty  0.5714286 0.4285714
  ##>   Van     0.5555556 0.4444444
``` 

#### Chi-squared test: Are row/column variables independent

Are `Type` and `Origin` independent:

``` r
chisq.test(Cars93$Type, Cars93$Origin)
``` 

#### 3 dimensional tables

``` r
ftable(Cars93$Man.trans.avail, Cars93$Origin, Cars93$Type)
  ##>              Compact Large Midsize Small Sporty Van
  ##>                                 
  ##> No  USA            2    11       9     0      0   4
  ##>     non-USA        0     0       4     0      0   2
  ##> Yes USA            5     0       1     7      8   1
  ##>     non-USA        9     0       8    14      6   2
``` 

### `base::table()` Examples

``` r
a0 = c("a", "b", "c", "a", "b", "a")
table(a0)
  ##> a b c
  ##> 3 2 1
str(table(a0))
  ##>  'table' int [1:3(1d)] 3 2 1
``` 

### `base::prop.table` Examples

It converts frequencies into fractions (percentages):

``` r
a0 = c("a", "b", "c", "a", "b", "a")
table(a0)
  ##> a b c
  ##> 3 2 1
prop.table(table(a0))
  ##>         a         b         c
  ##> 0.5000000 0.3333333 0.1666667
``` 

### read excel files

opt01: readxl read sheet by sheet

``` bash
readxl::read_excel("my-spreadsheet.xls", sheet = "data")
readxl::read_excel(path = sevk_emri_file, col_types = c("text", "date", "text", "text", "text", "text", "text", "text", "text", "numeric", "text", "text", "text", "text"))
``` 

### static web server

``` bash
Rscript -e 'servr::httd()' -p8000
``` 

### check version of a package

``` bash
packageVersion("stringr")
``` 

### arguments to Rscript  id=g_10954

Check `~/projects/study/r/ex/study_r/argument_to_rscript/e01.sh`

``` bash
Rscript e01.R some_arg
``` 

Check `~/projects/study/r/ex/study_r/argument_to_rscript/e01.R`

``` r
args <- commandArgs(trailingOnly = TRUE)
print(sprintf("args: %s", args[1]))
``` 

