---
title: "Study Statistics in R"
date: 2019-02-14T13:51:35+03:00 
draft: true
description: ""
tags:
categories: r statistics
type: post
url:
author: "Mert Nuhoglu"
output: rmarkdown::html_document
blog: mertnuhoglu.com
resource_files:
-
---

<style>
  .main-container {
    max-width: 1600px !important;
  }
</style>

``` {r set-options}
options(width = 150)
options(max.print = 30)
``` 

## Descriptive Statistics

``` r
library(tidyverse)
descriptive = function(df, vars) {
	df2 = df %>%
		dplyr::select(vars) 
	g01 = df2 %>%
		tidyr::gather("var", "value") 
	g02 = g01 %>%
		dplyr::count(var, value) %>%
		dplyr::group_by(var) %>%
		dplyr::filter(!is.na(value)) %>%
		dplyr::mutate(yuzde = prop.table(n) * 100) 
	s01 = g02 %>%
		dplyr::select(-n) %>%
		tidyr::spread(value, yuzde)
	gs = g01 %>%
		dplyr::group_by(var) %>%
		dplyr::summarize(
			ort = mean(value, na.rm=T)
			, sd = sd(value, na.rm=T)
			, eksik = sum(is.na(value))
			, n = dplyr::n() - eksik
		)
	s01 %>%
		dplyr::inner_join(gs, by = "var") %>%
		dplyr::arrange(var)
}
``` 

``` r
descriptive(dc, c("ws1","ws2","ws3","ws4","ws5","ws6"))
  ##>   var     `1`   `2`   `3`   `4`   `5`   ort    sd eksik     n
  ##>   <chr> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <int> <int>
  ##> 1 ws1   34.6  28.0  22.2   11.3  3.89  2.22  1.15     5   257
  ##> 2 ws2   13.4  21.3  30.3   27.2  7.87  2.95  1.16     8   254
  ##> 3 ws3   17.4  13.8  22.5   28.5 17.8   3.15  1.35     9   253
  ##> 4 ws4    3.92  5.10 16.1   40   34.9   3.97  1.03     7   255
  ##> 5 ws5    3.53  7.06  7.84  27.1 54.5   4.22  1.09     7   255
  ##> 6 ws6   19.6  11.5  20.8   21.2 26.9   3.24  1.46     2   260
``` 

## Article: A gentle introduction to Tidy statistics in R - Thomas Mock

Reading https://rpubs.com/jtmock/tidy-stats-in-R

Load libraries

```{r, message=F, warning = F, error = F}
# Load for ggplot, dplyr, tidyr, readxl
library(tidyverse)
library(readxl)
library(broom)
library(knitr)
```

Read excel file

```{r}
# read excel file
raw_df <- readxl::read_xlsx("ex/study_stats_in_R/ad_treatment.xlsx")

dplyr::glimpse(raw_df)
  ##> Observations: 600
  ##> Variables: 5
  ##> $ age            <dbl> 80, 85, 82, 80, 83, 79, 82, 79, 80, 79, 80, 79, 81, 83, 85, 81, 83, 85, 80, 80, 80, 85, 80, 83, 80, 82, 82, 84, 81, 81, 80, …
  ##> $ sex            <dbl> 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 1, 0, 1, 1, 0, 1, 0, 0, 1, 0, 0, 1, 0, 0, 0, 1, 0, 1, 1, 1, 1, 0, 1, 0, 0, 1, 0, 0, 0, 0, 1…
  ##> $ health_status  <chr> "Healthy", "Healthy", "Healthy", "Healthy", "Healthy", "Healthy", "Healthy", "Healthy", "Healthy", "Healthy", "Healthy", "He…
  ##> $ drug_treatment <chr> "Placebo", "Placebo", "Placebo", "Placebo", "Placebo", "Placebo", "Placebo", "Placebo", "Placebo", "Placebo", "Placebo", "Pl…
  ##> $ mmse           <dbl> 24.78988, 24.88192, 25.10903, 24.92636, 23.38845, 24.11390, 28.20327, 26.93959, 27.07407, 24.40350, 25.50462, 26.14665, 24.7…
```

Check distribution 

```{r}
# density plot

g2 <- ggplot(raw_df, aes(x = age)) +
  geom_density(fill = "blue")
g2
```

```{r}
# tidyverse way for range
raw_df %>% summarize(min = min(age),
                     max = max(age))
  ##>     min   max
  ##>   <dbl> <dbl>
  ##> 1    79    85
```

```{r}
# base way for range, note use of the $ to select column
range(raw_df$age)
  ##> [1] 79 85
```

Data cleaning:

```{r}
# check for NAs
raw_df %>% 
  summarize(na_count = sum(is.na(mmse)))
  ##>   na_count
  ##>      <int>
  ##> 1        0
```
  
Check experimental variables levels:

```{r}
# check Ns and levels for our variables
table(raw_df$drug_treatment, raw_df$health_status)
  ##>             Alzheimer's Healthy
  ##>   High Dose         100     100
  ##>   Low dose          100     100
  ##>   Placebo           100     100
table(raw_df$drug_treatment, raw_df$health_status, raw_df$sex)
  ##> , ,  = 0
  ##> 
  ##> 
  ##>             Alzheimer's Healthy
  ##>   High Dose          48      51
  ##>   Low dose           51      51
  ##>   Placebo            52      53
  ##> 
  ##> , ,  = 1
  ##> 
  ##> 
  ##>             Alzheimer's Healthy
  ##>   High Dose          52      49
  ##>   Low dose           49      49
  ##>   Placebo            48      47
```

Alternative way to check number of patients in each group:

```{r}
# tidy way of looking at variables
raw_df %>% 
  dplyr::group_by(drug_treatment, health_status, sex) %>% 
  dplyr::count()
  ##> # Groups:   drug_treatment, health_status, sex [12]
  ##>    drug_treatment health_status   sex     n
  ##>    <chr>          <chr>         <dbl> <int>
  ##>  1 High Dose      Alzheimer's       0    48
  ##>  2 High Dose      Alzheimer's       1    52
  ##>  3 High Dose      Healthy           0    51
  ##>  4 High Dose      Healthy           1    49
  ##>  5 Low dose       Alzheimer's       0    51
  ##>  6 Low dose       Alzheimer's       1    49
  ##>  7 Low dose       Healthy           0    51
  ##>  8 Low dose       Healthy           1    49
  ##>  9 Placebo        Alzheimer's       0    52
  ##> 10 Placebo        Alzheimer's       1    48
  ##> 11 Placebo        Healthy           0    53
  ##> 12 Placebo        Healthy           1    47
```

Visual exploratory data analysis:

```{r}
ggplot(data = raw_df, # add the data
       aes(x = drug_treatment, y = mmse, # set x, y coordinates
           color = drug_treatment)) +    # color by treatment
  geom_boxplot() +
  facet_grid(~health_status) # create panes base on health status
```

Summary statistics:

```{r}
sum_df <- raw_df %>% 
	dplyr::mutate(
		sex = factor(sex, labels = c("Male", "Female")),
		drug_treatment =  factor(drug_treatment, levels = c("Placebo", "Low dose", "High Dose")),
		health_status = factor(health_status, levels = c("Healthy", "Alzheimer's"))
	) %>% 
	dplyr::group_by(sex, health_status, drug_treatment ) %>%  # group by categorical variables
	dplyr::summarize(
		mmse_mean = mean(mmse),      # calc mean
		mmse_se = sd(mmse)/sqrt(dplyr::n()) # calc standard error
	) %>%  
	dplyr::ungroup() # ungrouping variable is a good habit to prevent errors
kable(sum_df)
  ##> |sex    |health_status |drug_treatment | mmse_mean|   mmse_se|
  ##> |:------|:-------------|:--------------|---------:|---------:|
  ##> |Male   |Healthy       |Placebo        |  24.93184| 0.2026394|
  ##> |Male   |Healthy       |Low dose       |  24.93751| 0.1961946|
  ##> |Male   |Healthy       |High Dose      |  25.70808| 0.1485297|
  ##> |Male   |Alzheimer's   |Placebo        |  12.12782| 0.2410401|
  ##> |Male   |Alzheimer's   |Low dose       |  14.65215| 0.2163392|
  ##> |Male   |Alzheimer's   |High Dose      |  23.13629| 0.1664349|
  ##> |Female |Healthy       |Placebo        |  24.94295| 0.2253879|
  ##> |Female |Healthy       |Low dose       |  24.97216| 0.2138110|
  ##> |Female |Healthy       |High Dose      |  25.36790| 0.1348688|
  ##> |Female |Alzheimer's   |Placebo        |  11.92192| 0.2565803|
  ##> |Female |Alzheimer's   |Low dose       |  15.32380| 0.1883008|
  ##> |Female |Alzheimer's   |High Dose      |  22.83295| 0.1611861|
```

```{r}
# save to excel
write.csv(sum_df, "ex/study_stats_in_R/adx37_sum_stats.csv")
```

Plotting summary statistics:

```{r}
g <- ggplot(data = sum_df, # add the data
       aes(x = drug_treatment,  #set x, y coordinates
           y = mmse_mean,
           group = drug_treatment,  # group by treatment
           color = drug_treatment)) +    # color by treatment
  geom_point(size = 3) + # set size of the dots
  facet_grid(sex~health_status) # create facets by sex and status
g
```

Statistics

```{r}
# set up the statistics df
stats_df <- raw_df %>% # start with data
   dplyr::mutate(drug_treatment = factor(drug_treatment, levels = c("Placebo", "Low dose", "High Dose")),
         sex = factor(sex, labels = c("Male", "Female")),
         health_status = factor(health_status, levels = c("Healthy", "Alzheimer's")))

dplyr::glimpse(stats_df)
  ##> Observations: 600
  ##> Variables: 5
  ##> $ age            <dbl> 80, 85, 82, 80, 83, 79, 82, 79, 80, 79, 80, 79, 81, 83…
  ##> $ sex            <fct> Male, Male, Male, Male, Male, Male, Male, Male, Female…
  ##> $ health_status  <fct> Healthy, Healthy, Healthy, Healthy, Healthy, Healthy, …
  ##> $ drug_treatment <fct> Placebo, Placebo, Placebo, Placebo, Placebo, Placebo, …
  ##> $ mmse           <dbl> 24.78988, 24.88192, 25.10903, 24.92636, 23.38845, 24.1…
```

Anova:

```{r}
# this gives main effects AND interactions
ad_aov <- aov(mmse ~ sex * drug_treatment * health_status, 
        data = stats_df)
```

```{r}
# look at effects and interactions
summary(ad_aov)
  ##>                                   Df Sum Sq Mean Sq  F value Pr(>F)
  ##> sex                                1      0       0    0.047  0.828
  ##> drug_treatment                     2   3601    1801  909.213 <2e-16 ***
  ##> health_status                      1  10789   10789 5447.953 <2e-16 ***
  ##> sex:drug_treatment                 2      8       4    2.070  0.127
  ##> sex:health_status                  1      5       5    2.448  0.118
  ##> drug_treatment:health_status       2   2842    1421  717.584 <2e-16 ***
  ##> sex:drug_treatment:health_status   2      5       2    1.213  0.298
  ##> Residuals                        588   1164       2
  ##> ---
  ##> Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
```

```{r}
# this extracts ANOVA output into a nice tidy dataframe
tidy_ad_aov <- tidy(ad_aov)
tidy_ad_aov
  ##> # A tibble: 8 x 6
  ##>   term                              df      sumsq    meansq statistic    p.value
  ##>   <chr>                          <dbl>      <dbl>     <dbl>     <dbl>      <dbl>
  ##> 1 sex                                1     0.0937   9.37e-2    0.0473  8.28e-  1
  ##> 2 drug_treatment                     2  3601.       1.80e+3  909.      1.18e-180
  ##> 3 health_status                      1 10789.       1.08e+4 5448.      1.57e-299
  ##> 4 sex:drug_treatment                 2     8.20     4.10e+0    2.07    1.27e-  1
  ##> 5 sex:health_status                  1     4.85     4.85e+0    2.45    1.18e-  1
  ##> 6 drug_treatment:health_status       2  2842.       1.42e+3  718.      1.67e-158
  ##> 7 sex:drug_treatment:health_sta…     2     4.80     2.40e+0    1.21    2.98e-  1
  ##> 8 Residuals                        588  1164.       1.98e+0   NA      NA

# which we can save to Excel
write.csv(tidy_ad_aov, "ex/study_stats_in_R/ad_aov.csv")
```

Pairwise unadjusted t tests:

```{r}
# pairwise t.tests
ad_pairwise <- pairwise.t.test(stats_df$mmse,
	stats_df$sex:stats_df$drug_treatment:stats_df$health_status, 
	p.adj = "none")
# look at the posthoc p.values in a tidy dataframe
kable(head(tidy(ad_pairwise)))
  ##> |group1                     |group2               |   p.value|
  ##> |:--------------------------|:--------------------|---------:|
  ##> |Male:Placebo:Alzheimer's   |Male:Placebo:Healthy | 0.0000000|
  ##> |Male:Low dose:Healthy      |Male:Placebo:Healthy | 0.9836212|
  ##> |Male:Low dose:Alzheimer's  |Male:Placebo:Healthy | 0.0000000|
  ##> |Male:High Dose:Healthy     |Male:Placebo:Healthy | 0.0050862|
  ##> |Male:High Dose:Alzheimer's |Male:Placebo:Healthy | 0.0000000|
  ##> |Female:Placebo:Healthy     |Male:Placebo:Healthy | 0.9685778|

# tidy
tidy_ad_pairwise <- tidy(ad_pairwise)

# save to excel
write.csv(tidy_ad_pairwise, "ex/study_stats_in_R/tidy_ad_pairwise.csv")
```

Pairwise Tukey post-hoc

```{r}
# call and tidy the tukey posthoc
tidy_ad_tukey <- tidy( TukeyHSD(ad_aov, 
	which = 'sex:drug_treatment:health_status')
)
write.csv(tidy_ad_tukey, "ex/study_stats_in_R/tukey_ad.csv")
kable(head(tidy_ad_tukey))
  ##> |term                             |comparison                                    |    estimate|    conf.low|   conf.high| adj.p.value|
  ##> |:--------------------------------|:---------------------------------------------|-----------:|-----------:|-----------:|-----------:|
  ##> |sex:drug_treatment:health_status |Female:Placebo:Healthy-Male:Placebo:Healthy   |   0.0111117|  -0.9140819|   0.9363052|   1.0000000|
  ##> |sex:drug_treatment:health_status |Male:Low dose:Healthy-Male:Placebo:Healthy    |   0.0056692|  -0.9000907|   0.9114291|   1.0000000|
  ##> |sex:drug_treatment:health_status |Female:Low dose:Healthy-Male:Placebo:Healthy  |   0.0403184|  -0.8748132|   0.9554500|   1.0000000|
  ##> |sex:drug_treatment:health_status |Male:High Dose:Healthy-Male:Placebo:Healthy   |   0.7762410|  -0.1295189|   1.6820009|   0.1775310|
  ##> |sex:drug_treatment:health_status |Female:High Dose:Healthy-Male:Placebo:Healthy |   0.4360609|  -0.4790707|   1.3511925|   0.9213646|
  ##> |sex:drug_treatment:health_status |Male:Placebo:Alzheimer's-Male:Placebo:Healthy | -12.8040144| -13.7053250| -11.9027038|   0.0000000|
```

```{r, eval = F}
write.csv(tidy_ad_pairwise, "ex/study_stats_in_R/tidy_ad_pairwise.csv")
write.csv(tidy_ad_tukey, "ex/study_stats_in_R/tidy_ad_tukey.csv")
```
  
Publication plot:

```{r}
sig_df <- tribble(
  ~drug_treatment, ~ health_status, ~sex, ~mmse_mean,
  "Low dose", "Alzheimer's", "Male", 17,
  "High Dose", "Alzheimer's", "Male", 25,
  "Low dose", "Alzheimer's", "Female", 18, 
  "High Dose", "Alzheimer's", "Female", 24
  )

sig_df <- sig_df %>% 
  dplyr::mutate(drug_treatment = factor(drug_treatment, levels = c("Placebo", "Low dose", "High Dose")),
         sex = factor(sex, levels = c("Male", "Female")),
         health_status = factor(health_status, levels = c("Healthy", "Alzheimer's")))
sig_df
  ##> # A tibble: 4 x 4
  ##>   drug_treatment health_status sex    mmse_mean
  ##>   <fct>          <fct>         <fct>      <dbl>
  ##> 1 Low dose       Alzheimer's   Male          17
  ##> 2 High Dose      Alzheimer's   Male          25
  ##> 3 Low dose       Alzheimer's   Female        18
  ##> 4 High Dose      Alzheimer's   Female        24
```

```{r}
# plot of cognitive function health and drug treatment
g1 <- ggplot(data = sum_df, 
       aes(x = drug_treatment, y = mmse_mean, fill = drug_treatment,  
           group = drug_treatment)) +
  geom_errorbar(aes(ymin = mmse_mean - mmse_se, 
                    ymax = mmse_mean + mmse_se), width = 0.5) +
  geom_bar(color = "black", stat = "identity", width = 0.7) +
  
  facet_grid(sex~health_status) +
  theme_bw() +
  scale_fill_manual(values = c("white", "grey", "black")) +
  theme(legend.position = "NULL",
        legend.title = element_blank(),
        axis.title = element_text(size = 20),
        legend.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 12)) +
  geom_text(data = sig_df, label = "*", size = 8) +
  labs(x = "\nDrug Treatment", 
       y = "Cognitive Function (MMSE)\n",
       caption = "\nFigure 1. Effect of novel drug treatment AD-x37 on cognitive function in healthy and demented elderly adults. \nn = 100/treatment group (total n = 600), * indicates significance at p < 0.001")
g1

```

```{r, eval = F}
# save the graph!
ggsave("ex/study_stats_in_R/ad_publication_graph.png", g1, height = 7, width = 8, units = "in")
```
