---
title: "Example: SQL Generation"
date: '`r strftime(Sys.time(), "%Y-%m-%dT%H:%M:%S+03:00")`'
draft: true
description: ""
tags:
categories: examples r sql
type: post
url:
author: "Mert Nuhoglu"
output: html_document
blog: mertnuhoglu.com
resource_files:
-
---

<style>
  .main-container {
    max-width: 1600px !important;
  }
</style>

``` {r set-options}
options(width = 250)
options(max.print = 30)
``` 

```{r}
library(dplyr, warn.conflicts = F)
```

# yumltordbschema::yuml_to_rdb()

``` r
yuml_to_rdb = function(data_model_dir) {
  cat_yuml.sh = system.file("bash/cat_yuml.sh", package = "yumltordbschema")
  system2(cat_yuml.sh, data_model_dir)

  build_yuml_data_model(data_model_dir)
  den = update_new_entities(data_model_dir)
  dfl = update_new_fields(den, data_model_dir)
  return(list(data_entity = den, data_field = dfl))
}
```

## yumltordbschema::build_yuml_data_model()


```{r}
lines = c("[enum_value| enum_value_id INT PK; enum_var_id INT FK; enum_value_name TEXT; ]",
  "[enum_var| enum_var_id INT PK; enum_var_name TEXT; ]",
  "[place| place_id INT PK; address TEXT; province TEXT; place_enum INT FK; ]")
lines
ydm = lines %>%
  stringr::str_replace_all( "^[ \\[]*", "" ) %>%
  stringr::str_replace_all( "[ \\]]*$", "" ) %>%
  stringr::str_replace_all( "\\|\\s*", "\\|" ) %>%
  stringr::str_replace_all( ";\\s*", ";" ) %>%
  dplyr::data_frame( ln = . ) %>%
  tidyr::separate( ln, c("entity_name", "columns"), "\\|" ) %>%
  tidyr::unnest( columns = strsplit( columns, ";") ) %>%
  tidyr::separate( columns, c("data_field_name", "type", "pk_fk"), " " ) %>%
  dplyr::mutate( pk_fk = ifelse( is.na(pk_fk), "NON_KEY", pk_fk)) %>%
  dplyr::mutate( type = ifelse( is.na(type), "TEXT", type)) %>%
  dplyr::mutate( entity_name = tolower(entity_name) ) %>%
  dplyr::mutate( data_field_name = tolower(data_field_name) ) 
ydm
``` 

## yumltordbschema::update_new_entities()

```{r}
den = ydm %>%
  dplyr::distinct( entity_name, .keep_all = T) %>%
  dplyr::select( entity_name ) %>%
  dplyr::mutate( data_entity_id = row_number() ) %>%
  dplyr::select( data_entity_id, dplyr::everything() )
den
```

## yumltordbschema::update_new_fields()

```{r}
dfl_p1 = ydm %>%
  dplyr::inner_join( den, by = "entity_name" ) %>%
  dplyr::select(-entity_name) %>%
  dplyr::mutate( data_field_id = row_number() ) %>%
  dplyr::select( data_field_id, dplyr::everything() )
dfl_p1
dfl_fk_id = dfl_p1 %>%
  dplyr::select( data_field_id, data_field_name, pk_fk ) %>%
  dplyr::filter( pk_fk == "FK" ) %>%
  dplyr::filter( rutils::greplm( data_field_name, "_id$") ) %>%
  dplyr::mutate( fk_data_entity_name = stringr::str_sub(data_field_name, end = -4)) %>%
  dplyr::select( data_field_id, fk_data_entity_name ) 
dfl_fk_id
dfl_fk_enum = dfl_p1 %>%
  dplyr::select( data_field_id, data_field_name, pk_fk ) %>%
  dplyr::filter( pk_fk == "FK" ) %>%
  dplyr::mutate( enum_var_name = data_field_name) %>%
  dplyr::filter( rutils::greplm( data_field_name, "_enum$") ) %>%
  dplyr::mutate( fk_data_entity_name = "enum_value") %>%
  dplyr::select( data_field_id, fk_data_entity_name, enum_var_name ) 
dfl_fk_enum
den_id = den %>%
  dplyr::select( data_entity_id, entity_name )
dfl_fk = dplyr::bind_rows(dfl_fk_id, dfl_fk_enum) %>%
  dplyr::left_join( den_id, by = c("fk_data_entity_name" = "entity_name") ) %>%
  dplyr::select(-fk_data_entity_name) %>%
  dplyr::select(data_field_id, fk_data_entity_id = data_entity_id, dplyr::everything()) 
dfl_fk
dfl = dfl_p1 %>%
  dplyr::left_join( dfl_fk, by = "data_field_id" )
dfl
```

Check enum fields:

```{r}
dfl_fk %>% dplyr::filter(!is.na(enum_var_name))
dfl %>% dplyr::filter(!is.na(enum_var_name))
```

# yumltordbschema::rdb_to_ddl()

Initial input arguments:

``` {r}
data_entity = den
data_field = dfl
```

``` {r}
den = data_entity 
dfl = data_field 
den
dfl
```

``` {r}
get_den_pk = function(den, dfl) {
  dfl_pk = dfl %>%
    dplyr::filter(pk_fk == "PK") %>%
    dplyr::select(-pk_fk) %>%
    dplyr::select(pk_data_field_id = data_field_id, fk_data_entity_id = data_entity_id, pk_data_field_name = data_field_name)
  den %>%
    dplyr::select(data_entity_id, entity_name) %>%
    dplyr::left_join( dfl_pk, by = c("data_entity_id" = "fk_data_entity_id")) %>%
    dplyr::select(data_entity_id, fk_data_entity_name = entity_name, fk_data_field_name = pk_data_field_name) 
}
den_pk = get_den_pk(den, dfl)
den_pk
``` 

``` {r}
get_dfl_fk = function(den, dfl) {
  dfl_fk = dfl %>%
    dplyr::filter(pk_fk == "FK") %>%
    dplyr::select(-pk_fk) %>%
    dplyr::left_join(get_den_pk(den, dfl), by = c("fk_data_entity_id" = "data_entity_id"))
}
dfl_fk = get_dfl_fk(den, dfl)
dfl_fk
``` 

Check if all fk fields are set properly:

``` {r}
dfl_fk %>% dplyr::filter(is.na(fk_data_entity_id))
``` 

Set up a `data.table` that will store `CREATE TABLE` sql statements:

``` {r}
sql = data.table::data.table(
                             data_entity_id = den$data_entity_id,
                             sql_create_table = ""
                             ) %>%
  data.table::setkey(data_entity_id)
ids = dplyr::distinct(dfl, data_entity_id)$data_entity_id
```

We will loop over each `data_entity_id` of all data fields `dfl`:

``` r
  for (deid in ids) {
    fd = dplyr::filter(dfl, data_entity_id == deid) %>%
      dplyr::left_join(den, by = "data_entity_id")
    table_name = unique(fd$entity_name)
    nonpks = dplyr::filter(fd, pk_fk != "PK")
    pks = dplyr::filter(fd, pk_fk == "PK")
    fks = dplyr::filter(fd, pk_fk == "FK") %>%
      dplyr::left_join(den_pk, by = "data_entity_id")
    sql[deid]$sql_create_table = create_table_sql( table_name, non_keys, pks, fks ) 
  }
```

Run the loop for `deid = 1` as an example:

``` {r}
deid = 1
fd = dplyr::filter(dfl, data_entity_id == deid) %>%
  dplyr::left_join(den, by = "data_entity_id")
fd
```

`fd` has the data fields of the current data entity among the ones we loop over:

``` {r}
table_name = unique(fd$entity_name)
non_keys = dplyr::filter(fd, pk_fk == "NON_KEY")
non_keys
pks = dplyr::filter(fd, pk_fk == "PK")
pks
fks = dplyr::filter(fd, pk_fk == "FK") %>%
  dplyr::left_join(den_pk, by = c( "fk_data_entity_id" = "data_entity_id"))
fks
```

Actual job of producing `CREATE TABLE` statements are done by `create_table_sql()` function:

Definition of `create_table_sql()` is here:

``` {r}
create_table_sql = function( table_name, non_keys, pks, fks ) {
  template_pk = "%s BIGINT GENERATED BY DEFAULT AS IDENTITY PRIMARY KEY"
  sql_pk = sprintf(template_pk, pks$data_field_name)
  template_fk = "%s int REFERENCES %s (%s)"
  sql_fk = sprintf(template_fk, fks$data_field_name, fks$fk_data_entity_name, fks$fk_data_field_name)
  template_nonkey = "%s %s"
  sql_nonkey = sprintf(template_nonkey, non_keys$data_field_name, non_keys$type)
  column_sql = c(sql_pk, sql_fk, sql_nonkey)
  column_names = column_sql %>% paste(collapse=", ")
  template = "CREATE TABLE %s (%s);"
  create_table = sprintf( template, table_name, column_names)
  return(create_table)
}
```

``` r
sql[deid]$sql_create_table = create_table_sql( table_name, non_keys, pks, fks )
``` 

## yumltordbschema::create_table_sql()

We have different templates for different types of data fields: PK fields, FK fields, non-key fields. These data fields are given as input to the function: `non_keys, pks, fks`

``` {r}
template_pk = "%s BIGINT GENERATED BY DEFAULT AS IDENTITY PRIMARY KEY"
sql_pk = sprintf(template_pk, pks$data_field_name)
sql_pk
```

``` {r}
template_fk = "%s int REFERENCES %s (%s)"
sql_fk = sprintf(template_fk, fks$data_field_name, fks$fk_data_entity_name, fks$fk_data_field_name)
sql_fk
```

``` {r}
template_nonkey = "%s %s"
sql_nonkey = sprintf(template_nonkey, non_keys$data_field_name, non_keys$type)
sql_nonkey
```

Now, we can concatenate all field definitions:

``` {r}
column_sql = c(sql_pk, sql_fk, sql_nonkey)
column_names = column_sql %>% paste(collapse=", ")
template = "CREATE TABLE %s (%s);"
create_table = sprintf( template, table_name, column_names)
create_table
``` 

Run the function at once:

``` {r}
create_table_sql( table_name, non_keys, pks, fks )
``` 

## Why Can't We Use `apply` instead of `for` loop?

In order to use `apply()`, we first need to wrap the body of `for` loop inside a function:

``` r
create_sql = function(fd) {
  table_name = unique(fd$entity_name)
  nonpks = dplyr::filter(fd, pk_fk != "PK")
  pks = dplyr::filter(fd, pk_fk == "PK")
  yumltordbschema::create_table_sql( 
    table_name, nonpks$data_field_name, nonpks$type, pks$data_field_name )
}
```

Now, we could have called the function like that:

``` r
apply(dfl, MARGIN = 1, create_sql)
```

But this will not work as expected, because in the above `for` loop we loop per each distinct `data_entity_aid`:

``` r
ids = dplyr::distinct(dfl, data_entity_aid)$data_entity_aid
for (deid in ids) {
  fd = dplyr::filter(dfl, data_entity_aid == deid)
  ...
}
```

In order to use `apply()` function, we need to ensure that each row of `dfl` contains a distinct `data_entity_aid`.

## Can I replace `for` loop with `lapply`?

Instead of `for` loop, I can use `lapply` here:

``` r
for (deid in ids) {
  fd = dplyr::filter(dfl, data_entity_id == deid) %>%
    dplyr::left_join(den, by = "data_entity_id")
  table_name = unique(fd$entity_name)
  nonpks = dplyr::filter(fd, pk_fk != "PK")
  pks = dplyr::filter(fd, pk_fk == "PK")
  fks = dplyr::filter(fd, pk_fk == "FK") %>%
    dplyr::left_join(den_pk, by = "data_entity_id")
  sql[deid]$sql_create_table = create_table_sql( 
    table_name, nonpks$data_field_name, nonpks$type, pks$data_field_name, fks )
}
```

It will be something like:

``` {r}
loop_sql_create_table = function(deid, dfl, den) {
  fd = dplyr::filter(dfl, data_entity_id == deid) %>%
    dplyr::left_join(den, by = "data_entity_id")
  table_name = unique(fd$entity_name)
  nonpks = dplyr::filter(fd, pk_fk != "PK")
  pks = dplyr::filter(fd, pk_fk == "PK")
  fks = dplyr::filter(fd, pk_fk == "FK") %>%
  dplyr::left_join(get_den_pk(den, dfl), by = "data_entity_id")
  return(tibble::tibble(
                        data_entity_id = deid
                        , sql_create_table = create_table_sql( table_name, non_keys, pks, fks )
                        )
         )
}
sql = lapply(ids, loop_sql_create_table, dfl, den) %>%
  dplyr::bind_rows()
sql
```

Which one is better?

I think `for` loop is easier to read and understand. But even so, I choose `lapply` over `for` loop because of composability. Having a side-effect-free function allows us to grow the program by composing building blocks.
